### 2013-02-13 doc/xps2013/proposal.tex
Incorporated Holt's suggestions. 	modified:   proposal.tex

\item \textbf{Language and compiler support:} Presently, applications are
hand-coded to the Grappa interface. A significant research thrust of the
proposed work is to ease this programming burden by language and compilation
support.

\item \textbf{Comprehensive benchmark suite:} Presently there does not exist a
comprehensive benchmark suite of irregular and graph applications. We are also
seeking funding through this proposal to put together just such a benchmark
suite. This suite will include benchmarks, reference implementations, and
input data-sets. In addition a website will be created \texttt{graphbench.org}
to distribute the benchmark suite and publish results.

@@

Grappa provides four methods for expressing parallelism. First, when the programmer identifies work that can be done in parallel, the work may be wrapped up in a function and queued with its arguments for later execution using a \texttt{spawn}. Second, a programmer can use \texttt{spawn\_on} to spawn a task on a specific core in the system or at the home core of a particular memory location. Third, the programmer can invoke a parallel for loop, provided that the trip count is known at loop entry. The programmer specifies a function pointer along with start and end indices and an optional threshold to control parallel overhead. Grappa does {\em recursive decomposition} of iterations, similar to Cilk's cilk\_for construct~\cite {cilkforimplementation} and TBB's {\tt parallel\_for}~\cite{intel_tbb}.  It generates a logarithmically-deep tree of tasks, stopping to execute the loop body when the number of iterations is below a specified threshold. Fourth, a programmer may want to execute an active message; that is, to run a small piece of code on a particular core in the system without waiting for execution resources to be available.  For example, custom synchronization primitives execute this way, as a function executed on the core where the data lives.

@@

\paragraph{Read and write combining: } When multiple remote memory operations are to be dispatched to the same node from the same thread, it is efficient to combine these together into a single request.  We will develop a compiler pass to combine operations when possible.

@@

@@

\paragraph{Task-weaving: } Much of the time the same task code is executing for hundreds of thousands of tasks in the Grappa runtime system.  For example, each task may be processing a loop body for a parallel for loop.  This is not always the most efficient approach to execution, and we already manually do recursive decomposition of parallel loops with a cut-off to group a collection of loop bodies together into a single task.  The task then executes each of these loop bodies sequentially.  This is not the most efficient approach, however.  An even more efficient technique is for the compiler to unroll and schedule all loop bodies simultaneously within the same task, thereby ``weaving'' these independent tasks together more efficiently to exploit computation / communication overlap.

@@

>>>>

\item \textbf{Language and compiler support:} Presently, Grappa is implemented
strictly as a library.  A significant research thrust of the proposed work is to use
language and compilation support to provide a cleaner programming model and use
statically-available knowledge to lessen the runtime overhead automatically.

\item \textbf{Comprehensive benchmark suite:} Presently there does not exist a
comprehensive benchmark suite of irregular and graph applications. We
are also seeking funding through this proposal to put together just
such a benchmark suite. This suite will include benchmarks, reference
implementations, and input data-sets. In addition a website will be
created to distribute freely the benchmark suite and publish results.

@@

Grappa provides two methods for expressing parallelism. First, when the programmer identifies work that can be done in parallel, the work may be wrapped up in a function and queued with its arguments for later execution using a \texttt{spawn}. A programmer can further specify using \texttt{spawn\_on} to spawn a task on a specific core in the system or at the home core of a particular memory location. Second, the programmer can invoke a parallel for loop, provided that the trip count is known at loop entry. The programmer specifies a function pointer along with start and end indices and an optional threshold to control parallel overhead. Grappa does {\em recursive decomposition} of iterations, similar to Cilk's cilk\_for construct~\cite {cilkforimplementation} and TBB's {\tt parallel\_for}~\cite{intel_tbb}.  It generates a logarithmically-deep tree of tasks, stopping to execute the loop body when the number of iterations is below a specified threshold.

@@

@@

\paragraph{Read and write combining: } As a special case of automatic delegate synthesis, when multiple remote memory operations are to be dispatched to the same node from the same thread, it may be more efficient to combine these together into a single request.  We will develop a compiler pass to combine operations when possible.

@@

\paragraph{Task-weaving: } Much of the time the same task code is executing for hundreds of thousands of tasks in the Grappa runtime system.  For example, each task may be processing a loop body for a parallel for loop.  However, tasks introduce overhead, so we already manually impose a cutoff on recursive decomposition of parallel loops.  This way, multiple loop bodies are collected into a single task and executed serially.  However, this limits the amount of parallelism available for latency tolerance. A mmore efficient technique is to allow the compiler to unroll and schedule all loop bodies simultaneously within the same task, including overlapping remote memory operations, thereby ``weaving'' these independent tasks together to overlap computation with communication with minimal task overhead.  Further, a set of ``woven tasks'' may be amenable to vector execution, allowing even more parallelism to be exploited.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%
%%%%% Extra stuff: collaboration plan and data-management plan
%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
